Minimizing post shock forecasting error using disparate information

### 2020-06-03 manuscript/manuscript.tex
changed Figure 2; made edits discussed in meeting

\t{\varepsilon}_{i} &\simiid  \normal{0}{\sigma^2_{\alpha}}, \qquad  i = 1, \ldots, n+1; \\

@@

In this example we forecast Conoco Phillips stock prices in the midst of the coronavirus recession. Specific interest is in predictions made after March 6th, 2020, the Friday before the stock market crash on March 9th, 2020. We will detail how we combine knowledge from disparate time series to improve the forecast of Conoco Phillips stock price that would be made without adjustment for the shock. The forecast consists of four steps: (1) pick a model, (2) selection of covariates,  (3) choices of donor pool, and (4) forecast. %Section \ref{background} provides the justification for the first three steps. Section \ref{results} discusses the result of our forecast.

@@

\begin{enumerate}
\item[(1)] {\bf AR(1) model and time window}. We will use a simple AR(1) model to forecast Conoco Phillips stock price. This model has been shown to beat no-change forecasts when predicting oil prices over time horizons of 1 and 3 months \citep{alquist2013forecasting}. We will consider 30 pre-shock trading days and we will forecast the immediate shock effect and the shock effect over a future five trading day window. All estimates will be adjusted for inflation. The model setup for AR(1) is exactly the same as what is stated in Section \ref{modelsetup} with addition of shock effects. All the parameters are estimated using OLS.
\item[(2)] {\bf Selection of covariates}. We perform our analyses incorporating daily S\&P 500 index prices and West Texas Intermediate (WTI) crude oil prices as covariates. 
\item[(3)] {\bf Construction of donor pool}. Our donor pool will consist of Conoco Phillips shock effects observed after September 11th, 2001, %March 18th, 2008, 
several events in September, 2008, and November 27, 2014. The first two shock effects were observed during recessions that were predicated by an inversion of the yield curve \citep{bauer2018economic}, and the third was an OPEC induced supply side shock effect \citep{huppmann2015opec}. The reasons for those three shocks are:
  \begin{enumerate}
  \item On September 11th, 2001, Islamic extremist al-Qaeda committed a series of terrorist attacks against the United States of America \citep{braniff2011towards}. In addition to the tragic loss of life, this triggered negative repercussions on the U.S. and world economies with falling stock prices and a falling U.S. dollar \citep{floyd2001terror}. Moreover, trading was closed until September 17, 2001 \citep{johnston2006impact}.
  \item  In early September 2008, time series of oil prices experienced a sudden increase in volatility simultaneously due to turmoil in financial markets. The political, economic, social or environmental events may coincide with these shocks \citep{ewing2013volatility}. Notable shock effects followed the placement of Fannie May and Freddie Mac in conservatorship on September 7th (shock effect on the 8th), Lehman Brothers filing for bankruptcy on September 15th, and the Office of Thrift Supervision closes Washington Mutual Bank on September 25th \citep{dwyer2009financial, longstaff2010subprime}.
  \item On November 27th, 2014, it is documented that oil prices fall as OPEC opts not to cut production \citep{huppmann2015opec}. During the Great Recession when economic activity clearly declined, both oil and stock prices fell which points to demand factors. During the second half of 2014, oil prices plummeted but equity prices generally increased, suggesting that supply factors were the key driver \citep[Page 19]{baffes2015great}. 
  \end{enumerate}
\item[(4)] {\bf Forecast}. \textbf{Jilei, please fill this in. I do not know what you have in mind.}
\end{enumerate}

@@

{\bf Jilei, can you add the weighted adjustment estimator to this analysis?}
We estimate the five shock effects mentioned above using OLS and then construct $\hat{\alpha}_{\text{adj}}$ and $\hat{\alpha}_{\text{IVW}}$ from these estimates. We also consider estimation without the September 11th, 2001 shock effect since there was a week long delay in trading due to the terrorist attack. We also study additive shock effect estimators where the shock effects corresponding to separate supply and demand shocks are added to estimate the unknown shock effect. The supply shock donor pool consists of the November 27th, 2014 shock effect and the demand shock donor pool consists of the remaining shock effects.

@@

We can see from Figure~\ref{Fig:CP} that $\hat{\alpha}_{\text{adj}}$ and $\hat{\alpha}_{\text{IVW}}$ perform decently well, they do not recover the magnitude of the shock effect but are much better than unadjusted forecasts that do not account for shock effects.
The unadjusted forecast had an RMSE of 12.24, 
the use of $\hat{\alpha}_{\text{adj}}$ had an RMSE of 6.85,
the use of $\hat{\alpha}_{\text{adj}}$ without September 11th had an RMSE of 5.96,
the use of $\hat{\alpha}_{\text{IVW}}$ had an RMSE of 9.37,
and the use of $\hat{\alpha}_{\text{IVW}}$ without September 11th had an RMSE of 6.15.
The additive adjustment estimator computed by adding the $\hat{\alpha}_{\text{adj}}$ estimators for the demand and supply shock effects without September 11th has an RMSE of 0.80. 
The additive adjustment estimator computed by adding the $\hat{\alpha}_{\text{IVW}}$ estimators for the demand and supply shock effects without September 11th has an RMSE of 0.023.

We also see that additive estimators do extremely well. However, these additive estimators are retrospective but their is some apriori justification for their use. \citet{kilian2009not} study the effect that different supply and demand shocks have on oil prices through a vector auto regressive model. Their model postulates an additive nature of shock effects, although the additivity parameters requires estimation in their context. %This provides some anecdotal justification for apriori estimation of the March 9th Conoco Phillips shock effect with the addition of the adjustment estimator corresponding to the separate supply and demand aspects of this shock effect.

>>>>

\t{\varepsilon}_{i} &\simiid  \E{\t{\varepsilon}}=0, \var{\t{\varepsilon}}=\sigma^2_{\alpha}, \qquad i = 1, \ldots, n+1,\\

@@

In this example we forecast Conoco Phillips stock prices in the midst of the coronavirus recession. Specific interest is in predictions made after March 6th, 2020, the Friday before the stock market crash on March 9th, 2020. We will detail how we combine knowledge from disparate time series to improve the forecast of Conoco Phillips stock price that would be made without adjustment for the shock. The forecast consists of four steps: (1) pick a model, (2) selection of covariates,  (3) choices of donor pool. %Section \ref{background} provides the justification for the first three steps. Section \ref{results} discusses the result of our forecast.

@@

\begin{enumerate}
\item[(1)] {\bf AR(1) model and time window}. We will use a simple AR(1) model to forecast Conoco Phillips stock price. This model has been shown to beat no-change forecasts when predicting oil prices over time horizons of 1 and 3 months \citep{alquist2013forecasting}. We will consider 30 pre-shock trading days and we will forecast the immediate shock effect and the shock effect over a future five trading day window. All estimates will be adjusted for inflation. The model setup for AR(1) is exactly the same as what is stated in Section \ref{modelsetup} with addition of shock effects. All the parameters are estimated using OLS.
\item[(2)] {\bf Selection of covariates}. We perform our analyses incorporating daily S\&P 500 index prices and West Texas Intermediate (WTI) crude oil prices as covariates. 
\item[(3)] {\bf Construction of donor pool}. Our donor pool will consist of Conoco Phillips shock effects observed after September 11th, 2001, %March 18th, 2008, 
several events in September, 2008, and November 27, 2014. The first two shock effects were observed during recessions that were predicated by an inversion of the yield curve \citep{bauer2018economic}, and the third was an OPEC induced supply side shock effect \citep{huppmann2015opec}. The reasons for those three shocks are:
  \begin{enumerate}
  \item On September 11th, 2001, Islamic extremist al-Qaeda committed a series of terrorist attacks against the United States of America \citep{braniff2011towards}. In addition to the tragic loss of life, this triggered negative repercussions on the U.S. and world economies with falling stock prices and a falling U.S. dollar \citep{floyd2001terror}. Moreover, trading was closed until September 17, 2001 \citep{johnston2006impact}.
  \item  In early September 2008, time series of oil prices experienced a sudden increase in volatility simultaneously due to turmoil in financial markets. The political, economic, social or environmental events may coincide with these shocks \citep{ewing2013volatility}. Notable shock effects followed the placement of Fannie May and Freddie Mac in conservatorship on September 7th (shock effect on the 8th), Lehman Brothers filing for bankruptcy on September 15th, and the Office of Thrift Supervision closes Washington Mutual Bank on September 25th \citep{dwyer2009financial, longstaff2010subprime}.
  \item On November 27th, 2014, it is documented that oil prices fall as OPEC opts not to cut production \citep{huppmann2015opec}. During the Great Recession when economic activity clearly declined, both oil and stock prices fell which points to demand factors. During the second half of 2014, oil prices plummeted but equity prices generally increased, suggesting that supply factors were the key driver \citep[Page 19]{baffes2015great}. 
  \end{enumerate}
%\item[(4)] {\bf Forecast}. \textbf{Jilei, please fill this in. I do not know what you have in mind.}
\end{enumerate}

@@

{\bf Jilei, can you add the weighted adjustment estimator to this analysis?}
We estimate the five shock effects mentioned above using OLS and then construct $\hat{\alpha}_{\text{adj}}$ and $\hat{\alpha}_{\text{IVW}}$ from these estimates. We also consider estimation without the September 11th, 2001 shock effect since there was a week long delay in trading due to the terrorist attack.

@@

We can see from Figure~\ref{Fig:CP} that $\hat{\alpha}_{\text{adj}}$ and $\hat{\alpha}_{\text{IVW}}$ perform decently well, they do not recover the magnitude of the shock effect but are much better than unadjusted forecasts that do not account for shock effects.
The unadjusted forecast had an RMSE of 12.24, 
the use of $\hat{\alpha}_{\text{adj}}$ had an RMSE of 6.85,
the use of $\hat{\alpha}_{\text{adj}}$ without September 11th had an RMSE of 5.96,
the use of $\hat{\alpha}_{\text{IVW}}$ had an RMSE of 9.37,
and the use of $\hat{\alpha}_{\text{IVW}}$ without September 11th had an RMSE of 6.15.

We also study additive shock effect estimators where the shock effects corresponding to separate supply and demand shocks are added to estimate the unknown shock effect. The supply shock donor pool consists of the November 27th, 2014 shock effect and the demand shock donor pool consists of the remaining shock effects. The additive adjustment estimator computed by adding the $\hat{\alpha}_{\text{adj}}$ estimators for the demand and supply shock effects without September 11th has an RMSE of 0.80.  The additive adjustment estimator computed by adding the $\hat{\alpha}_{\text{IVW}}$ estimators for the demand and supply shock effects without September 11th has an RMSE of 0.023.

These additive adjustment estimators do extremely well, nearly perfectly forecasting the realized shock effect. There is apriori justification for the use of these simply additive adjustment estimators, although their nearly perfect performance in this example is a retrospective finding. Our apriori justification comes from \citet{kilian2009not} who study the effect that different supply and demand shocks have on oil prices through a vector auto regressive model. Their model postulates an additive nature of shock effects, although the additivity parameters requires estimation in their context. %This provides some anecdotal justification for apriori estimation of the March 9th Conoco Phillips shock effect with the addition of the adjustment estimator corresponding to the separate supply and demand aspects of this shock effect.
